---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyr)
library(readr)
library(dplyr)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

```{r}
deaths <- av |>
  pivot_longer(cols = c("Death1", "Death2", "Death3", "Death4", "Death5"),
               names_to = "Time",
               values_to = "Death") |>
  filter(Death != "")

deaths$Time <- parse_number(deaths$Time)

head(deaths)
```

Similarly, deal with the returns of characters.

```{r}
returns <- av |>
  pivot_longer(cols = c("Return1", "Return2", "Return3", "Return4", "Return5"),
               names_to = "Time",
               values_to = "Return") |>
  filter(Return != "")

returns$Time <- parse_number(returns$Time) 

head(returns)
```
Based on these datasets calculate the average number of deaths an Avenger suffers. 

```{r}
avg_deaths <- deaths |>
  group_by(Name.Alias) |>
  summarise(Total_Deaths = sum(Death == "YES"), .groups = "drop") |>
  summarise(Avg_Deaths = mean(Total_Deaths))

avg_deaths
```

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.


Kate Passwater:
Quote: Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team"
Code:
```{r}
total_avengers <- length(av$Name.Alias)
total_avengers
total_died_once <- deaths |> filter(Time == 1) |> filter(Death == "YES") |> summarise(diedOnce=n())
total_died_once
```
Summary for Kate's Results: This quote is correct. There are 173 Avengers in this table and 69 of them have died at least once.


Samara Feldhacker:

Quote: 

> I counted 89 total deaths — some unlucky Avengers are basically Meat Loaf with an E-ZPass — and on 57 occasions the individual made a comeback

Code: 

```{r}
deaths |>
  filter(Death == "YES") |>
  summarise(totalDeaths = n())
```

```{r}
returns |>
  filter(Return == "YES") |>
  summarise(comebackcount = n()) 
```

Summary for Samara's results: The statement is true. In the code shown above, it is shown that the total number of Deaths is 89 and the total number of returns is 57.


Shashank Koneru:  
Quote: "But you can only tempt death so many times. There’s a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife, but only a 50 percent chance they recovered from a second or third death."  
Code: 
```{r}
initial_stats <- av |>
  summarise(
    initial_deaths = sum(Death1 == "YES", na.rm = TRUE),
    initial_returns = sum(Return1 == "YES", na.rm = TRUE),
    initial_return_rate = initial_returns / initial_deaths
  )

two_third_stats <- av |>
  summarise(
    two_third_deaths = sum(Death2 == "YES", na.rm = TRUE) + sum(Death3 == "YES", na.rm = TRUE),
    two_third_returns = sum(Return2 == "YES", na.rm = TRUE) + sum(Return3 == "YES", na.rm = TRUE),
    two_third_return_rate = two_third_returns / two_third_deaths
  )

initial_stats
two_third_stats
```
Summary for Shashank's Results: This statement is backed up by the data and the results. Out of the 69 initial deaths, 46 were followed by a return, which provided a return rate of 0.667 (roughly 2 in 3). For the second and third deaths combined, there were 18 deaths followed by 9 returns. This is a return rate of 50%, making the article's claim accurate.



Sebastian Naranjo:
Quote: 
Code: 




### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.
